---
title: "Data Visualisation in R"
author: "Tinus Cloete"
date: "December 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

## Introduction

This project is focused on chapter 3 of the R for data scientists book by Hadley Wickam. I am also focusing on learning to use R markdown and git at the same time. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

Remember that this is how to bold text: **Bold**

## Data Visualisation

Data Exploration is the art of looking at your data, rapidly generating hypotheses, quickly testing them and then repeating the process again and again.

The goal is to generate many possible leads that can later then be explored in more depth.

You begin by asking a question. 
It's important to try to answer it yourself, and to be precise. e.g. what is the relationship between the variables (instead of just saying yes and no)

Lets start with the Fuel Economy data set: mpg

These are the details about the data set.

Fuel economy data from 1999 and 2008 for 38 popular models of car

This dataset contains a subset of the fuel economy data that the EPA makes available on http://fueleconomy.gov. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car.

```{r mpg_data}
#The ? function doesn't work in markdown.
#?mpg

#Fuel economy data from 1999 and 2008 for 38 popular models of car

summary(mpg)
length(unique(mpg$cty))
```

The question is: Do cars with bigger engines use more fuel than cars with smaller engines? What is the relationship?

I expect it to be to be true and a negative linear relationship (bigger engine, lower fuel economy).  

***

#Introduction to GGPlot

Important variables:

cty: city miles per gallon  
displ: engine displacement, in liters  
hwy: highway miles per gallon  

Focusing on the highway miles per gallon for now.

```{r plot1_mpg}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y= hwy))
```

The relationship is negative. But there are a few cars that have an unexpected high range for their engine size. Could these cars be hybrids or something?

```{r plot2_mpg}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y= hwy, color = class))
```

By differentiating the plot by class, we can see that these cars are sport cars. This makes sense, due to the smaller size of the cars.

This plot is all blue... it is outside the aesthetic.

```{r plot3_mpg}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y= hwy), color="blue")
```

```{r plot4_mpg}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y= hwy))+
  facet_wrap(~ class, nrow=2)
```

```{r plot5_mpg, include=FALSE}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y= hwy))+
  facet_wrap(~ cty)
```

```{r smooth_plot}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```


```{r smooth_plot2}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r multi_plot_1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y= hwy, color = drv))+
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv, linetype = drv))
```

```{r multi_plot_2}
ggplot(data = mpg, mapping = aes(x=displ, y= hwy)) + 
  geom_point(mapping=aes(color = class))+
  geom_smooth()
```

```{r multi_plot_3}
ggplot(data = mpg, mapping = aes(x=displ, y= hwy)) + 
  geom_point(mapping=aes(color = class))+
  geom_smooth(data=filter(mpg,class=="subcompact"),se=FALSE)
```

#Diamonds and Bar charts

Prices of 50,000 round cut diamonds

Description

A dataset containing the prices and other attributes of almost 54,000 diamonds. The variables are as follows:

A data frame with 53940 rows and 10 variables:

price:price in US dollars (\$326–\$18,823)  
carat:weight of the diamond (0.2–5.01)  
cut:quality of the cut (Fair, Good, Very Good, Premium, Ideal)  
color:diamond colour, from J (worst) to D (best)  

clarity:a measurement of how clear the diamond is (I1 (worst), SI1,   SI2, VS1, VS2, VVS1, VVS2, IF (best))  
  
x:length in mm (0–10.74)  
y:width in mm (0–58.9)  

z:depth in mm (0–31.8)  

depth:total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43–79)

table:width of top of diamond relative to widest point (43–95)

```{r bar1}

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))

```

```{r bar2}
ggplot(data=diamonds)+
  stat_count(mapping=aes(x=cut))
```


```{r override_stat}

demo <- tribble(
  ~a, ~b,
  "bar1", 10,
  "bar2", 20,
  "bar3", 30
)

ggplot(data=demo)+
  geom_bar(mapping=aes(x=a,y=b),stat="identity")

```

```{r override_calc_var}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,y=..prop..,group=1))
```

```{r play_with_stats}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,y=price),stat="sum")

```

```{r play_with_stats2}


demo <- tribble(
  ~a, ~b,
  "bar1", 10,
  "bar2", 20,
  "bar3", 30,
  "bar3", 40.5252
)

ggplot(data=demo)+
#  geom_bar(mapping=aes(x=a),stat="identity")
   geom_col(mapping=aes(x=a,y=b))

```

```{r stat_summ_plot}

ggplot(data=diamonds)+
  stat_summary(mapping=aes(x=cut,y=depth),
               fun.ymin=min,
               fun.ymax=max,
               fun.y=median)


```

```{r bar_fill1}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=cut))
```

```{r bar_fill2}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity))
```

```{r bar_position_1}
#position identity. All would overlap. More useful for 
#point plots. To see the overlap have to set the 
#transparency (alpha) or fill = NA (completely transparent)

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),alpha=1/5,
           position="identity")

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,color=clarity),position="identity",fill=NA)

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="fill")

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=clarity,fill=cut),position="dodge")

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="dodge")


```

#Coordinate Systems

```{r coordinateflip}
ggplot(data=mpg,mapping=aes(x=class,y=hwy))+
  geom_boxplot()+
  coord_flip()
```

```{r map_nz}
nz<-map_data("nz")

#??? group=group??

ggplot(nz, aes(long, lat, group=group))+
  geom_polygon(fill="white",colour="black")+
  coord_quickmap()
  
```

```{r polar}

bar <- ggplot(data=diamonds)+
       geom_bar(mapping=aes(x=cut,fill=cut),
           show.legend=FALSE,
           width=1)+
       theme(aspect.ratio=1)+
       labs(x=NULL,y=NULL)
  
bar+coord_flip()
bar+coord_polar()

```

```{r polar_pie}
 
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=1,fill=cut),
           position="stack",
           width=1)+
  labs(x=NULL,y=NULL)+
  theme(aspect.ratio=1)+
  coord_polar(theta="y")

```

